\section{Exact algorithms}

During the internship, we developed two exact algorithms. The first one is time efficient but has a huge memory cost, and the second one is memory efficient but is quite time consuming.

\subsection{A time efficient exact algorithm}

We first show a preliminary version of our algorithm: Algorithm \ref{algo:ckcom1} which works by doing a single pass over the $k$-cliques and by storing all $(k-1)$-cliques that are contained in a $k$-clique. It relies on the fact that the output of CPM is a partition of the $k$-cliques of the input graph such that two 
$k$-cliques are in a same partition's cluster if they share $(k-1)$-nodes: a partition's cluster is thus a connected component in the graph where nodes are $k$-cliques and two $k$-cliques are connected if they share $(k-1)$ nodes.

Thus to compute all $k$-clique communities it suffice to list all $k$-cliques while merging the communities of two $k$-cliques when they 
share a $(k-1)$-clique.

\vspace{0.3cm}

\noindent {\bf Implementation issues.} Algorithm \ref{algo:ckcom1} can be implemented efficiently by listing all $k$-cliques in the input graph using efficient algorithm such as \cite{chiba1985arboricity}; storing all $(k-1)$-cliques (contained in a $k$-clique) and using a Union-Find data structure\footnote{\url{https://en.wikipedia.org/wiki/Disjoint-set_data_structure}.}. Given a Union-Find data structure $\tt UF$ containing $n$ elements, the following two operations are allowed.
\begin{itemize}
\item ${\tt UF.Find}(a)$ returns the partition's cluster ID of element $a$ in $O(\alpha(n))$ time where $\alpha$ is the inverse Ackermann function which is essentially constant.
\item ${\tt UF.Union}(p,q)$ merges the two clusters $p$ and $q$ in $O(\alpha(n))$ time and returns the ID of the resulting cluster.
{\color{red} As $p$ and $q$ are not
elements, they are clusters IDs, so
should Union complexity be $O(1)$ ? }

\item ${\tt UF.MakeSet}(a)$ creates a new cluster and assigns $a$ to it while returning the cluster ID in constant time.
\end{itemize}
In order to have a more concise pseudocode, we also define ${\tt UF.FindOrCreate}(a)$ which returns the partition's cluster ID of element $a$ and if $a$ is not in $\tt{UF}$ it creates a new cluster and assigns $a$ to it while returning the cluster ID.

%In addition for each cluster in the partition of the $(k-1)$-cliques, we also store the set of nodes in that cluster, these overlapping clusters of nodes is the output of our algorithm.
\vspace{0.3cm}

\begin{algorithm}[!htbp]
\caption{One pass over $k$-cliques, storing $(k-1)$-cliques}
\label{algo:ckcom1}
\begin{algorithmic}[1]
  \State ${\tt UF} \leftarrow$ Union-Find datastructure 
    \For {each $k$-clique $c_k \in G$}
        \State $p \leftarrow NULL$
    	\For {each $(k-1)$-clique $c_{k-1}$ in $c_k$}
      		\State $q\leftarrow {\tt UF.FindOrCreate} (c_{k-1})$
          	\State $p \leftarrow {\tt UF.Union}(p,q) $\Comment{Returns $q$ if $p=NULL$}
      	\EndFor
    \EndFor
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{algo:ckcom1} needs to store all $(k-1)$-cliques (contained in a $k$-clique) of the input graph. As the number of such $(k-1)$-cliques can be very large, this algorithm is problematic in some cases as it uses too much memory.

We next show how to avoid storing all these $(k-1)$-cliques. In particular, instead of storing the communities each $k$-clique belongs to, we store the communities each edge belongs to and we use a new datastructure that we call {\it Overlapping Union-Find} (OUF). This new datastruture builds on top of Union-Find and allows to do the same operations considering overlapping sets rather than a partition (i.e. non-overlapping sets).

We first detail how our datastructure works and then depict two pseudocodes:
\begin{itemize}
\item Algorithm \ref{algo:ckcom2} is very similar to Algorithm \ref{algo:ckcom1}, but is using an OUF instead of a UF. This 
\todo{which ? Algo1 ?}
algorithm creates a variable for each $(k-1)$-cliques found and is thus not efficient in terms of memory like Algorithm \ref{algo:ckcom2}. We show it for illustration purpose only.
\item Algorithm \ref{algo:ckcom3} is our final algorithm also using an OUF. This algorithm does not create a variable for each $(k-1)$-cliques and is much more efficient than our two first algorithms.
\end{itemize}

As we will see in Section~\ref{sec:analysis} and in Section~\ref{sec:experiments}, in addition of being orders of magnitude faster than the state-of-the-art, our final algorithm (Algorithm~\ref{algo:ckcom3}) is very efficient in terms of memory consumption. This allows our program to solve the problem for unprecedented values of $k$ and sizes of graphs using a commodity machine.

\vspace{0.3cm}

{\color{red}
OUF is a known data structure ?
If not we should explain in more details,
especially prove stuff about the complexity.
}
\noindent {\bf Overlapping Union-Find.} Given an Overlapping Union-Find data structure $\tt OUF$, we denote by $n$ the number of overlapping clusters in $\tt OUF$ and given an element 
$a$, $n_a$ denotes the number of clusters $a$ belongs to. 
The following operations are allowed.
\begin{itemize}
\item ${\tt OUF.Find}(a)$ returns the set of clusters that element $a$ belongs to in time $O(n_a\cdot\alpha(n))$;
\item $OUF.Union(p,q)$ merges the two clusters $p$ and $q$ in time $O(\alpha(n))$;
{\color{red} As $p$ and $q$ are not
elements, they are clusters IDs, so
should Union complexity be $O(1)$ ? }


\item $OUF.MakeSet()$ creates an empty cluster and returns it in time $O(1)$ and
\item $OUF.Add(a,p)$ add element $a$ to cluster $p$ (do nothing if $a$ is already in $p$) in time $O(1)$.
\end{itemize}

\todo{These operations are allowed by storing variables representing the overlapping clusters in a Union-Find datastructure and by pointing each element to the clusters that it belongs to.}

In order to have a clearer and more concise pseudocodes, we generalize the operations on sets of elements or clusters. In particular, the following operations are allowed.\\
Used in Algorithm \ref{algo:ckcom2}:
 \begin{itemize}
  \item ${\tt OUF.FindOrCreate}(A)$ returns the intersection of the sets of clusters of the elements in $A$: (i) if the intersection is a singleton, it returns a single cluster; (ii) if the intersection is empty, it creates a new cluster and add all elements in $A$ to it while returning its ID.
 \end{itemize}
Used in Algorithm \ref{algo:ckcom3}:
 \begin{itemize}
  \item ${\tt OUF.Find}(A)$ returns the intersection of the sets of clusters of the elements in $A$: (i) it returns a single cluster if the intersection is a singleton; (ii) it returns $NULL$ if the intersession is empty;
  
{  \color{red}
Hi, I propose to simplify and say just that 
it "returns the intersection of the sets of clusters of the elements in $A$." replacing lines 5-7 by only one line

        $S \leftarrow S \cup {\tt OUF.Find}(E(c_{k-1}))$
}
  \item $OUF.UnionOrCreate(P)$ merges all clusters in $P$ into a single one and returns the ID of the resulting cluster, if $P$ is empty it creates a new empty cluster and returns its ID.
  \item $OUF.Add(A,p)$ add all elements in $A$ to cluster $p$ (do nothing for an element already in $p$).
\end{itemize}

Note that following this description of operations, in line 5 of Algorithm~\ref{algo:ckcom2}, 
${\tt OUF.FindOrCreate}(E(c_{k-1}))$ always returns a single cluster and in line 5 of Algorithm~\ref{algo:ckcom2}, ${\tt OUF.Find}(E(c_{k-1}))$ always returns a single cluster or $NULL$ 
as the set of all edges belonging to a given $(k-1)$-clique can only belong to a single community.
\vspace{0.3cm}

\begin{algorithm}[!htbp]
\caption{One pass over $k$-cliques, storing edges}
\label{algo:ckcom2}
\begin{algorithmic}[1]
  \State ${\tt OUF} \leftarrow$ Overlapping Union-Find datastructure 
    \For {each $k$-clique $c_k \in G$}
        \State $q \leftarrow NULL$
    	\For {each $(k-1)$-clique $c_{k-1}$ in $c_k$}
      		\State $p \leftarrow {\tt OUF.FindOrCreate} (E(c_{k-1}))$
          	\State $q \leftarrow {\tt OUF.Union} (p,q) $ \Comment{Returns $q$ if $p=NULL$}
      	\EndFor
\EndFor
\end{algorithmic}

\end{algorithm}

\begin{algorithm}[!htbp]
\caption{One pass over $k$-cliques, storing edges}
\label{algo:ckcom3}
\begin{algorithmic}[1]
\State $OUF \leftarrow$ Overlapping Union-Find datastructure 
\For {each $k$-clique $c_k \in G$}
	\State $S\leftarrow \emptyset$
	\For {each $(k-1)$-clique $c_{k-1}$ in $c_k$}
        \State $p \leftarrow {\tt OUF.Find}(E(c_{k-1}))$%\Comment{setID containing all edges in $E$}
        \If {$p\neq NULL$}
        	\State $S \leftarrow S\cup \{p\}$
        \EndIf
    \EndFor
    \State $q\leftarrow {\tt OUF.UnionOrCreate}(S)$
	\State ${\tt OUF.Add}(E(c_k),q)$ 
\EndFor
\end{algorithmic}
\end{algorithm}


{\color{red} 
Algo 2 and 3 produce exactly the same result !
}
%Note that for $k=3$, 

\input{Contents/gr0.tex}

\input{Contents/gr1.tex}

We now proceed to the theoretical analysis of our algorithms.



\subsection{A memory efficient exact algorithm}

A community is a set of $k$-cliques. Keeping in memory all its $k$-cliques, or all its $(k-1)$-cliques, is not scaling well for huge graphs. A way of tackling this issue is to find a way to compress the data. In this section, the main idea is to store a community as a graph, storing only the nodes and the edges of the $k$-cliques added. So, we say that a $k$-clique is in the community if all its nodes and edges are in the community. Consequently, it can happen that a $k$-clique is tested beeing in the community whereas it has not be added to it. So, the community appears to have more $k$-clique that should be. The main issue is that if a $k$-clique appears to be in a community, and is also in another one, the two communities are merged, whereas it should not.

Therefore, whereas keeping the community as a list of edges allows to store communities with few memory, it creates parasites cliques that we need to compute and to store. A $k$-clique is added to the community if it shares a $(k-1)$-clique with it. A $(-1)$-clique parasite of a community is a $(k-1)$-clique that is tested to be in the community but which is not part of any $k$-clique added to the community. In consequence, if the $(k-1)$-clique common to a $k$-clique and a community is a parasite, we need to know it not to add the $k$-clique to the community.

In consequence, we choose to use a procedure to list all the $(k-1)$-cliques parasites of a community. Every time an edge  is added to the community, we list all the $(k-1)$-cliques of the community containing this edge. Each one which is not a sub-clique of the $k$-clique added is then a parasite $(k-1)$-clique (RESULTAT A METTRE EN VALEUR ? + PREUVE ?). Then, we remove all the $(k-1)$-subliques of the $k$-clique added, so the list of $(k-1)$-cliques parasites is always up-to-date.

\begin{algorithm}
  \caption{CPM 3 condensé}
  \begin{algorithmic}[1]
    \State GraphCommus $\gets dict()$ % \Comment{Graphs of communities}
    \State GhostCliques $\gets dict()$ % \Comment{Ghost cliques of communities}
    \State Fusions $\gets []$ % \Comment{Fusions to perform at the end of the algorithm}
    \For{each $k$-clique $c_k \in G$}
    \State commus\_ck $\gets$ \Call{CommusNeighboursOf}{$c_k$, GraphCommus, GhostCliques}
    %\State
    \If{commus\_ck == $\emptyset$} %\Comment{No current community is a neighbour of $c_k$}
    \State \Call{CreateCommu}{$c_k$, GraphCommus, GhostCliques} \Comment{$c_k$ : new community}
    %\State
    \Else \Comment{Pick one neighbour community and add $c_k$ to it, computing new ghost cliques for each new edge}
    \State $commu$ $\gets$ ONE community from commus\_ck
    \State \Call{AddToCommuGraph}{$c_k$, GraphCommus, GhostCliques}
    \If{|commus\_ck| > 1} \Comment{Record the other communities to merge them at the end}
    \State Fusions.push(commus\_ck) 
    \EndIf
    \EndIf
    \EndFor
    %\State 
    \State \Call{ApplyFusions}{Fusions, GraphCommus}
  \end{algorithmic}
\end{algorithm}